---
title: "Project #1: Grade Calculator"
output: html_document
---

******
## 1. Project Details

### Goal of the project
Calculates the final grade given the weight of each hwk, midterm, final exam and lab report.

The freshmen physics courses at Hunter College of The City University of New York are split into Lecture and Labs. Lecture accounts for 85% of the final grade and labs for 15%. The labs and lecture are taught by different instructors. Combine the data by id number to calculate the final grade.

### The lab grades
* The last two columns represent quiz 1 and quiz 2. Each quiz is graded on the scale of 0-20. Each quiz is worth 10% of the lab grade.
* Columns 1-12 are grades from lab reports. They are graded in the scale of 0-100. The lowest two should be dropped when calculating the final lab grade, take the best 10 out of 12. The lab reports are worth 80% of the lab grade. The students must have minimum of 10 labs to recieve a passing grade.

### The lecture grades
* There are two midterms and a final:
   + The lowest midterm is worth 20% and the highest 35%. 
   + The final exam is worth 45%.

******
## 2. Load and clean data

Load necessary libraries, readODS for reading ODS data, xlsx for reading xlsx spreadsheet, dplyr for data wrangling, stringr for string operations.
```{r, message=FALSE}
library(readODS)
library(xlsx)
library(dplyr)
library(stringr)
```

First, load the lecture grades from xlsx spread sheet into a data frame named lecture. Change student id from "id" to "ID", elimimlate the while space in the student id.

```{r}
lecture <- read.xlsx2("/Users/qiu/Downloads/DSBootcamp/project1_data/lecture.xlsx",
                     sheetName = "Sheet1",
                     colIndex=c(2:5))
names(lecture) <- names(lecture) %>% str_replace_all("id", "ID") # change from “id” to “ID”,
lecture <- mutate(lecture, ID = str_replace_all(ID, "[[ s]]", ""))  # remove letter "s" and the white space in student id's.
lecture
```

Second, load the lab grades from xlsx spread sheet for lab section 1L12.
Notice there is a typo in the name of collumn "lab5", which shoulb be "Lab6" instead. Make corrections to the name in the code.

```{r}
lab120sec12 <- read.xlsx("/Users/qiu/Downloads/DSBootcamp/project1_data/Fall 2015 Lab 120, Sec 1L12.xlsx", 
                   sheetName = "Sheet1", 
                   rowIndex = c(8:20), 
                   colIndex = c(1:15))
names(lab120sec12) <- names(lab120sec12) %>% 
  str_replace_all("\\.","") %>%
  str_replace("lab5","LAB6") %>%
  str_replace("StudentsName","ID")
lab120sec12 <- mutate(lab120sec12, ID = str_replace_all(ID, "s", ""))  # remove letter "s" in student id's.
lab120sec12
```

Last, load the lab grades from ods spread sheet for lab section 1L09. The same typo in the name of collumn "lab5", which shoulb be "Lab6" instead.

```{r}
ods1 <- read.ods("/Users/qiu/Downloads/DSBootcamp/project1_data/Fall 2015 Lab 120 section 1L09.ods",
                 sheet = 1)
# View(lab120sec09)
lab120sec09 = data.frame(ods1[9:28,1:15])
names(lab120sec09) = c(ods1[8,1:13],ods1[30,14:15])

names(lab120sec09) <- names(lab120sec09) %>% 
  str_replace("lab 5","LAB6") %>%
  str_replace("Student's Name","ID") %>%
  str_replace_all("\\#","") %>%
  str_replace_all(" ", "") %>%
  toupper()
lab120sec09 <- mutate(lab120sec09, ID = str_replace_all(ID, "S", "")) 
lab120sec09
```

## 3. Process data

Use rbind funcion to join two grades in two separate lab sections vertically, then use merge function to joint lecture and lab grades horizontally.

```{r}
lab <- rbind(lab120sec09, lab120sec12)
class_grade <- merge(lecture,lab,by="ID")
class_grade
```
## 4. Analysis data


